Reasoning About a Machine with Local Capabilities

### 2019-03-06 toplas/paper.tex
[TOPLAS] Make minor changes to ML text

Let us first look at this ML program.
At the top-level, it is a lambda function that can be invoked by the context.
When it is invoked, it allocates a fresh mutable variable $x$ of integer type, initially containing the value $0$.
Next, the function returns a second closure (let's call this $\textit{cl}_x$) that can be invoked by the context whenever it chooses.
$\textit{cl}_x$ itself takes a callback function \texttt{adv} that it will invoke twice after setting $x$ to $0$ and $1$ respectively.
After the second invocation of \texttt{adv}, $\textit{cl}_x$ will verify that $x$ is still set to $1$.

This last assertion, at line 8 of Figure~\ref{fig:awkward}, should never fail.
From the code, this seems natural, since \texttt{adv} does not have access to $x$.
Therefore, after the second invocation at line 7, $x$ should still be in the state it was before that invocation.
However, consider that \texttt{adv} may have access to $\textit{cl}_x$ and within that second invocation of \texttt{adv}, it may invoke $\textit{cl}_x$ again.
Then $\textit{cl}_x$ would set $x$ to $0$ again (on line 5) and reinvoke another callback.
If that second callback were somehow able to skip its own caller and return directly to the caller of \texttt{adv}, it would end up at line 8 of Figure~\ref{fig:awkward} with $x$ set to $1$, causing the assertion to fail.
Without going into details, such an attack is perfectly possible if \texttt{adv} has access to a call/cc primitive (or equivalent).
In other words, if we are able to prove that the assertion will never fail, this attack, and other similar attacks, on well-bracketed control flow are adequately prevented.

@@

The attack that we explained above is actually more natural at this low-level.
The callback \texttt{adv} now gets a return capability.
As explained, it could invoke $\mathit{cl}_x$ again, during its second invocation, with a second callback function $\texttt{adv}'$ for $\mathit{cl}_x$ to invoke.
If \texttt{adv} had some way of passing its return capability to $\texttt{adv}'$ (by storing it somewhere on the stack or in the heap, or hide it an unused register), then the assertion could be made to fail.
However, our calling convention prevents any of this from happening, as we prove in the following lemma.

>>>>

Let us first look at this ML program.
At the top-level, it is a lambda function that can be invoked by the context.
When it is invoked, it allocates a fresh mutable variable $x$ of integer type that initially contains the value $0$.
Next, the function returns a second closure (let's call this $\textit{cl}_x$) that can be invoked by the context whenever it chooses.
$\textit{cl}_x$ itself takes a callback function \texttt{adv} that it will invoke twice after setting $x$ to $0$ and $1$ respectively.
After the second invocation of \texttt{adv}, $\textit{cl}_x$ will verify that $x$ is still set to $1$.

The assertion on line 8 of Figure~\ref{fig:awkward} should never fail.
From the code, this seems natural since \texttt{adv} does not have access to $x$.
Therefore after the second invocation on line 7, $x$ should still be in the state it was before that invocation.
However, \texttt{adv} could have access to $\textit{cl}_x$ and could invoke $\textit{cl}_x$ again within the second invocation of \texttt{adv}.
Then $\textit{cl}_x$ would set $x$ to $0$ again (on line 5) and reinvoke another callback.
If that second callback were somehow able to skip its own caller and return directly to the caller of \texttt{adv}, it would end up on line 8 of Figure~\ref{fig:awkward} with $x$ set to $1$, causing the assertion to fail.
Without going into details, such an attack is perfectly possible if \texttt{adv} has access to a call/cc primitive (or equivalent).
In other words, if we are able to prove that the assertion will never fail, this attack, and other similar attacks, on well-bracketed control flow are adequately prevented.

@@

The attack that we explained above is actually more natural at this low-level.
The callback \texttt{adv} now gets a return capability.
As explained, it could invoke $\mathit{cl}_x$ again, during its second invocation, with a second callback function $\texttt{adv}'$ for $\mathit{cl}_x$ to invoke.
If \texttt{adv} had some way of passing its return capability to $\texttt{adv}'$ (by storing it on the stack or in the heap, or hide it in an unused register), then the assertion could be made to fail.
However, our calling convention prevents any of this from happening as we prove in the following lemma.
